---
title: "scCTS Package Vignette"
package: scCTS
output:
  BiocStyle::html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
vignette: >
  %\VignetteIndexEntry{scCTS Package Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r knitr-options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(scCTS)
```


# Input of `scCTS`
`scCTS` is built on top of the Bioconductor's `SingleCellExperiment` class, which is a commonly used data structure for storing scRNA-seq datasets. In the package we provide a toy simulated dataset stored in a `SingleCellExperiment` object. You can directly load it through the code:
```{r}
data("sim.sce")
sim.sce
```
The dataset `sim.sce` has 2,000 genes and 10,000 cells. There are two columns in the 
`colData` slot, named as `celltype` and `subject`. The `celltype` column should store
the cell type labels of each cells, and the `subject` column should store the subject labels of each cells. Although `scCTS` requires predefined cell type labels, you can first cluster the cells in your own dataset and then use the cluster labels if your dataset has no cell annotations.


# Parameters of `scCTS`
```{r}
scCTS.res <- scCTS(sim.sce,use.raw = TRUE, subject.rep='subject', celltype.rep='celltype', numCores=2, effect_thres=0, tol=1e-5)
```
You can run `scCTS` in simply one line of code. `use.raw = TRUE` means `scCTS` will
extract the raw counts from the `sim.sce` and normalize the raw counts internally. If
`use.raw = FALSE` and no assay name is specified by `use.norm.rep`, `scCTS` will try
to extract the normalized counts by calling `normcounts()` on the input `SingleCellExperiment` object. `subject.rep` and `celltype.rep` specify the names of
columns storing the cell type and subject information. `numCores` is number of cores used for parallel computation. `effect_thres` is a threshold for filtering genes with negative mean. `tol` is the tolerance in the EM estimation.

# Results of `scCTS`
For each unique cell type, `scCTS` creates a list to store the corresponding estimated parameters and posterior probabilities for each gene to be cell-type specific. For example, we show some posterior probabilities of genes for `celltype1`:
```{r}
head(scCTS.res$celltype1$pp.d1)
```


# Run Classic DE Methods
Except from our proposed method, this package also provides a common interface for classic DE methods such as the Wilcoxon test, t-test and [`DESeq2`](https://bioconductor.org/packages/release/bioc/html/DESeq2.html):
```{r}
bs.res <- runBaselineMethod(sim.sce,use.raw=TRUE, per.subject=TRUE,  
                            celltype.rep='celltype', subject.rep='subject', 
                            method='wilcox', numCores=2)
```
The parameter `method` specifies which method to use. `per.subject=TRUE` means the   method will run separately on each subject. If you set `per.subject=FALSE` then the
method will ignore the subject information and run on all cells. Other parameters are the same as `scCTS`.


# Results of Classic DE Methods
Results of all classic methods share a common structure. Each type of information
(statistics, P-values, adjusted P-values, etc.) is stored in a 3 dimensional array.
The first dimension is for genes, the second dimension is for cell types, and the 
third dimension is for subjects. If you set `per.subject=FALSE`, then the name of the
last dimension is `"all"`.
```{r}
head(bs.res$wilcox.fdr_info[,,'subject1'])
```


# Session Info
```{r}
sessionInfo()
```
